---
title: "Doug State Section"
author: "Doug Walton"
date: "May 3, 2016"
output: html_document
---

This is my contribution to the State section.



First, to get set up
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(readr)
library(broom)
library(knitr)


```
#### notes form Team Project
1. upload state wide data; explain data cleaning process both manual and R. 

2. do a graph for each demographic and crime rate for ALL states for the most current year. 
3. single out top 4 and bottom 4 states with worst crime rate and show a chart of them.
5. show graphs using facet_wrap for top 4 states and the demographics data for current year, for over time; then for bottom 4;
6. any correlations between the demographics for top 4 states? for bottom 4 states? do comparisons between the top 4 states to see trends. do you notice any trends between top 4 states data vs bottom 4's data in terms of which demographics jump out at you?


# State Demographic Data
State data for 2012 compiled from multiple  American Community Survey tables (US Census Bureau) and from the FBI Uniform Crime Reporting database.


```{r}

stdem <- read.csv("states-key-demographics-2012.csv")


```

```{r}
# Population by state - choropleth view
# inspired by - See more at: http://www.trulia.com/blog/tech/the-choroplethr-package-for-r/
# can be installed from CRAN
# USCensus API key needed install one time. To obtain: http://api.census.gov/data/key_signup.html to get the key
# then api.key.install(key="592bc14cnotarealkey686552b17fda3c89dd389") (example, done one time)
# see http://eglenn.scripts.mit.edu/citystate/wp-content/uploads/2013/06/wpid-working_with_acs_R3.pdf for acs package info.

library(choroplethr)
library(acs)
library(choroplethrMaps)
data("state.map")

# state population graphic
# extract and set to expectation of choropleth package
sf <- stdem %>% select(State,tot_pop)
colnames(sf) <- c("region","value")  
sf$region <- tolower(sf$region)

state_choropleth(sf, title = '2012 U.S State Population')

# SAMEEN - this can be removed, but I left in case it might help the story
# rape Pct Graphic
#df <- read.csv("state_rape_percent_2009.csv")
#state_choropleth(df, title = 'Rape Pct 2009')

# median age graphic
#cf <- read.csv("Sources/state-media-age-2010.csv")
#state_choropleth(cf, title = 'Median Age 2010')

#Note: You can see intuitively flipping between the maps that there could be a correspondence of age and rape pct. 

# although a minor negative correlation? (rape pct down as median age up) Size of certain male age groups would be better probably

#qf <- left_join(df,cf, by = 'region')
#qf %>% ggplot(aes(x = value.x)) + geom_point(aes(y= value.y))
#cor.test(qf$value.x,qf$value.y)

```



```{r}

# Alernative view of population by state - graph view
stbypop <- stdem %>% arrange(desc(tot_pop)) 
stbypop$State <- factor(stbypop$State, levels = stbypop$State)
stbypop %>% ggplot(aes(x= State, y = tot_pop/100000)) + geom_bar(stat = "identity") + coord_flip()


```



```{r}
# PUt in the percent for Murder and Rape  by State
stcrime <-stdem
stcrime <- stcrime %>% mutate(rape_pct = num_rapes/tot_pop * 100000, murder_pct = num_murder/tot_pop * 100000)
stcrime <- stcrime %>% arrange(rape_pct) 
stcrime$State <- factor(stcrime$State, levels = stcrime$State)

# Make the graph for rape and murder - would be nice if clustered bar
stcrime %>% ggplot(aes(x= State, y = rape_pct, fill = murder_pct)) + geom_bar(position='dodge',stat = "identity") + coord_flip() + labs(title = "Rates of Rape and Murder by State", y = "Number Rapes per 100K people")


```
The bars of the graph show the rape percent, and the color shows the murder percent. As the blue colors are fairly widely distributed, there doesn't seem to be a high correspondence of murder to rape.

The plot also shows us the four worst and best states, listed below, which we will use later.

```{r}
# worst states
ws <- stcrime %>% arrange(desc(rape_pct)) %>% select(State, rape_pct)

kable(head(ws, n=4), digits = 2, col.names = c("State","Rapes per 100K"), caption = "Highest Number of Rapes")

# best states
bs <- stcrime %>% select(State, rape_pct)
kable(head(bs, n=4), digits = 2, col.names = c("State","Rapes per 100K"), caption = "Lowest Number of Rapes")

```


In this next part, we explore a number of factors that could contribute to a high incidence of rape. A similar format is used to compare the factor against the rate of rape. 

```{r}

# Median age - bar version
stcrime %>% ggplot(aes(x= State, y = rape_pct, fill = median_age)) + geom_bar(position='dodge',stat = "identity") + coord_flip() + labs(title = "Rates of Rape and Median Age by State", y = "Number Rapes per 100K people")

# Median age - colored state version
stcrime %>% ggplot(aes(x= median_age, y = rape_pct, fill = State, color=State, cex = 2)) + geom_point() + labs(title = "Rates of Rape and Median Age by State", y = "Number Rapes per 100K people")

# Median age - lm smoothed line (using colored states also wouldnt work)
stcrime %>% ggplot(aes(x= median_age, y = rape_pct)) + geom_point() + labs(title = "Rates of Rape and Median Age by State", y = "Number Rapes per 100K people") + geom_smooth(method = lm)

# Median age - loess smoothed line
stcrime %>% ggplot(aes(x= median_age, y = rape_pct)) + geom_point() + labs(title = "Rates of Rape and Median Age by State", y = "Number Rapes per 100K people") + geom_smooth()

# Median Income
stcrime %>% ggplot(aes(x= median_inc, y = rape_pct)) + geom_point() + labs(title = "Rates of Rape and Average Income by State", y = "Number Rapes per 100K people") + geom_smooth()

# Educational attainment vs rape
# This gives a comparison factor for states by adding percentages for educational attainment for over 24 year olds, where the categories are less than grade 9, somewhere between 9 and 12, and high school grad
stedu <- stcrime %>% mutate(non_hs = pct_less_9_over_24+pct_some_hs_over_24)

# bar version
stedu %>% ggplot(aes(x= State, y = rape_pct, fill = non_hs)) + 
geom_bar(position='dodge',stat = "identity") + coord_flip()  + labs(title = "Rates of Rape and High School Completion by State", y = "Number Rapes per 100K people")

# smoothed version
stedu %>% ggplot(aes(x= non_hs, y = rape_pct)) + 
geom_point() + labs(title = "Rates of Rape and High School Completion by State", y = "Number Rapes per 100K people", x = "Percent of people over 24 without HS completion") + geom_smooth()

# HS completion
# bar version
stedu %>% ggplot(aes(x= State, y = rape_pct, fill = pct_hs_over_24)) + geom_bar(position='dodge',stat = "identity") + coord_flip()  + labs(title = "Rates of Rape and High School Completion by State", y = "Number Rapes per 100K people")

# smooth version
stedu %>% ggplot(aes(x= pct_hs_over_24, y = rape_pct)) + 
geom_point() + labs(title = "Rates of Rape and High School Completion by State", y = "Number Rapes per 100K people") + geom_smooth()

```

On these, it looks like median age could have an effect. Otherwise, these don't seem to have much impact. 

So, we trying a few more things. 

```{r}
# Percent of males - more males equals more prospective rapists
# bar version
stcrime %>% ggplot(aes(x= State, y = rape_pct, fill = pct_male)) + geom_bar(position='dodge',stat = "identity") + coord_flip()  + labs(title = "Rates of Rape and Percen Males by State", y = "Number Rapes per 100K people")

# smoothed version - seems to show modest rise with more education!
stcrime %>% ggplot(aes(x= pct_male, y = rape_pct)) + 
geom_point() + labs(title = "Rates of Rape and Percent of Males by State", y = "Number Rapes per 100K people", x = "Percent Males") + geom_smooth()

# Young Males - younger age group may be more inclined
# bar version
stcrime$pct_young_male <- stcrime$pct_young_male * 100
stcrime %>% ggplot(aes(x= State, y = rape_pct, fill = pct_young_male)) + geom_bar(position='dodge',stat = "identity") + coord_flip()  + labs(title = "Rates of Rape and Younger Males by State", y = "Number Rapes per 100K people")

# smoothed version
stcrime %>% ggplot(aes(x= pct_young_male, y = rape_pct)) + 
geom_point() + labs(title = "Rates of Rape and Age of Males by State", y = "Number Rapes per 100K people", x = "Percent Young Males") + geom_smooth()

# Racial Mix using % white as proxy 
# bar version
stcrime %>% ggplot(aes(x= State, y = rape_pct, fill = pct_white)) + geom_bar(position='dodge',stat = "identity") + coord_flip()   + labs(title = "Rates of Rape and Percent Whites by State", y = "Number Rapes per 100K people")

# smoothed version
stcrime %>% ggplot(aes(x= pct_white, y = rape_pct)) + 
geom_point() + labs(title = "Rates of Rape and Percentage Whites by State", y = "Number Rapes per 100K people", x = "Percent White") + geom_smooth()

# Population density - larger, more open spaces with less oversight
# bar version. filter out DC which throws it off.
stcrime_lessDC <- stcrime %>% filter(State != "District of Columbia")
stcrime_lessDC %>% ggplot(aes(x= State, y = rape_pct, fill = pop_density)) + geom_bar(position='dodge',stat = "identity") + coord_flip()  + labs(title = "Rates of Rape and Population Density by State", y = "Number Rapes per 100K people")

#smoothed version
stcrime_lessDC %>% ggplot(aes(x= pop_density, y = rape_pct)) + 
geom_point() + labs(title = "Rates of Rape and Population Density by State", y = "Number Rapes per 100K people", x = "Population Density") + geom_smooth()

# smooth version focusing on the lower end of pop density
stcrime_lessDC %>% filter(pop_density <= 250) %>% ggplot(aes(x= pop_density, y = rape_pct)) + geom_point() + labs(title = "Rates of Rape and Population Density by State", y = "Number Rapes per 100K people", x = "Population Density") + geom_smooth()

```

Of those, perhaps the following seemed to have some correlation:

Percent of males - More males in general seems correlated with higher incidence of rape.

Median age - as discussed above, lower median age seems a little related to higher incidence of rape.

Population Density - it does seem as if at least in states where the population density is less than 100 people per square mile there is a significantly higher incidence of rape


```{r}
# Let's look at some of the factors on those specifically

worst <- stcrime %>% filter(State %in% c("Alaska","South Dakota","Michigan","New Mexico")) %>% select(State,pct_male,pop_density,median_age)
worst$pct_male <- as.numeric(as.character(worst$pct_male))

# Collect this for later
wsum <- worst %>% summarise(mean(pct_male), mean(pop_density), mean(median_age))

# Transform for facets
worst <- worst %>% gather(key = variable, Value = pct_male:median_age)

# Other ways to do it
#ggplot(worst, aes(x=State, y=value)) + geom_bar(stat='identity') + facet_grid(.~variable, scales = "free") +coord_flip() + labs(x='',y='')

#ggplot(worst,aes(x=State, y=value)) + geom_bar(stat='identity') + facet_grid(variable~., scales = "free") +coord_flip() + labs(x='',y='')

ggplot(worst,aes(x=State, y=value)) + geom_bar(stat='identity') + facet_grid(variable~., scales = "free") + labs(title = "Worst States in 2012", x='',y='')


```


And here are the four states with the best percentage, compared.

```{r}
best <- stcrime %>% filter(State %in% c("New Jersey","New York","Virginia","Vermont")) %>% select(State,pct_male,pop_density,median_age)
best$pct_male <- as.numeric(as.character(best$pct_male))

# Collect for later
bsum <- best %>% summarise(mean(pct_male), mean(pop_density), mean(median_age))

# Transform for facets
best <- best %>% gather(key = variable, Value = pct_male:median_age)

ggplot(best,aes(x=State, y=value)) + geom_bar(stat='identity') + facet_grid(variable~., scales = "free")  + labs(title = "Best States in 2012", x='',y='')


```

So, there are some possible indicators viewed this way. Racial mix, in the end seems, unlikely. However, larger, less policed areas with less people may be a contributing factor, and a greater percentage of males may increase the odds of one being prone to misconduct. A perhaps quick way to summarize them is below. 

```{r}
# make a table of averages
s <- c("Worst","Best")
sumt <- rbind(wsum,bsum)
sum <- data.frame(s,sumt)
kable(sum)

```

