---
title: "Doug State Section"
author: "Doug Walton"
date: "May 3, 2016"
output: html_document
---

This is my contribution to the State section.



First, to get set up
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(readr)
library(broom)
library(knitr)


```
#### notes form Team Project
1. upload state wide data; explain data cleaning process both manual and R. 

2. do a graph for each demographic and crime rate for ALL states for the most current year. 
3. single out top 4 and bottom 4 states with worst crime rate and show a chart of them.
5. show graphs using facet_wrap for top 4 states and the demographics data for current year, for over time; then for bottom 4;
6. any correlations between the demographics for top 4 states? for bottom 4 states? do comparisons between the top 4 states to see trends. do you notice any trends between top 4 states data vs bottom 4's data in terms of which demographics jump out at you?


# State Demographic Data
State data for 2012 compiled from multiple  American Community Survey tables (US Census Bureau) and from the FBI Uniform Crime Reporting database.


```{r}

stdem <- read.csv("states-key-demographics-2012.csv")

# Population by state
stbypop <- stdem %>% arrange(desc(tot_pop)) 
stbypop$State <- factor(stbypop$State, levels = stbypop$State)
stbypop %>% ggplot(aes(x= State, y = tot_pop/100000)) + geom_bar(stat = "identity") + coord_flip()

# PUt in the percent for Murder and Rape  by State
stcrime <-stdem
stcrime <- stcrime %>% mutate(rape_pct = num_rapes/tot_pop * 100000, murder_pct = num_murder/tot_pop * 100000)
stcrime <- stcrime %>% arrange(desc(rape_pct)) 
stcrime$State <- factor(stcrime$State, levels = stcrime$State)

# Make the graph for rape and murder
stcrime %>% ggplot(aes(x= State, y = rape_pct, fill = murder_pct)) + geom_bar(position='dodge',stat = "identity") + coord_flip() + labs(title = "Rates of Rape and Murder by State", x = "Number Rapes per 100,000")


```
The bars of the graph show the rape percent, and the color shows the murder percent. As the blue colors are fairly widely distributed, there doesn't seem to be a high correspondence of murder to rape.

The plot also shows us the four worst and best states, listed below, which we will use later.

Four with highest rape percent
Alaska
South Dakota
Michigan
New Mexico

Four with lowest rape percent
New Jersey
New York
Virginia
Vermont

In this next part, we explore a number of factors that could contribute to a high incidence of rape. A similar format is used to compare the factor against the rate of rape. 

```{r}

# Median age
stcrime %>% ggplot(aes(x= State, y = rape_pct, fill = median_age)) + geom_bar(position='dodge',stat = "identity") + coord_flip() + labs(title = "Rates of Rape and Median Age by State", x = "Number Rapes per 100,000")

# Average Income
stcrime %>% ggplot(aes(x= State, y = rape_pct, fill = median_inc)) + geom_bar(position='dodge',stat = "identity") + coord_flip() + labs(title = "Rates of Rape and Average Income by State", x = "Number Rapes per 100,000")

# Educational attainment vs rape
# This gives a comparison factor for states by adding percentages for educational attainment for over 24 year olds, where the categories are less than grade 9, somewhere between 9 and 12, and high school grad
stedu <- stcrime %>% mutate(non_hs = pct_less_9_over_24+pct_some_hs_over_24)
stedu %>% ggplot(aes(x= State, y = rape_pct, fill = non_hs)) + geom_bar(position='dodge',stat = "identity") + coord_flip()  + labs(title = "Rates of Rape and High School Completion by State", x = "Number Rapes per 100,000")

# HS completion
stedu %>% ggplot(aes(x= State, y = rape_pct, fill = pct_hs_over_24)) + geom_bar(position='dodge',stat = "identity") + coord_flip()  + labs(title = "Rates of Rape and High School Completion by State", x = "Number Rapes per 100,000")


```

These don't seem to have much. If anything, it looks a little like lower HS attainment is related to lower rates of rape. 

So, we trying a few more things. 

```{r}
# Percent of males - more males equals more prospective rapists
stcrime %>% ggplot(aes(x= State, y = rape_pct, fill = pct_male)) + geom_bar(position='dodge',stat = "identity") + coord_flip()

# Young Males - younger age group may be more inclined
stcrime$pct_young_male <- stcrime$pct_young_male * 100
stcrime %>% ggplot(aes(x= State, y = rape_pct, fill = pct_young_male)) + geom_bar(position='dodge',stat = "identity") + coord_flip()

# Racial Mix using % white as proxy 
stcrime %>% ggplot(aes(x= State, y = rape_pct, fill = pct_white)) + geom_bar(position='dodge',stat = "identity") + coord_flip()

# Population density - larger, more open spaces with less oversight
stcrime_lessDC <- stcrime %>% filter(State != "District of Columbia")
stcrime_lessDC %>% ggplot(aes(x= State, y = rape_pct, fill = pop_density)) + geom_bar(position='dodge',stat = "identity") + coord_flip()



```

Of those, perhaps the following seemed to have some correlation:
Percent of males
Percent whites
Population Density


```{r}
# Let's look at some of the factors on those specifically

worst <- stcrime %>% filter(State %in% c("Alaska","South Dakota","Michigan","New Mexico")) %>% select(State,pct_male,pop_density,pct_white)
worst$pct_male <- as.numeric(as.character(worst$pct_male))

# Collect this for later
wsum <- worst %>% summarise(mean(pct_male), mean(pop_density), mean(pct_white))

# Transform for facets
worst <- worst %>% gather(key = variable, Value = pct_male:pct_white)

# Other ways to do it
#ggplot(worst, aes(x=State, y=value)) + geom_bar(stat='identity') + facet_grid(.~variable, scales = "free") +coord_flip() + labs(x='',y='')

#ggplot(worst,aes(x=State, y=value)) + geom_bar(stat='identity') + facet_grid(variable~., scales = "free") +coord_flip() + labs(x='',y='')

ggplot(worst,aes(x=State, y=value)) + geom_bar(stat='identity') + facet_grid(variable~., scales = "free") + labs(title = "Worst States in 2012", x='',y='')


```


And here are the four states with the best percentage, compared.

```{r}
best <- stcrime %>% filter(State %in% c("New Jersey","New York","Virginia","Vermont")) %>% select(State,pct_male,pop_density,pct_white)
best$pct_male <- as.numeric(as.character(best$pct_male))

# Collect for later
bsum <- best %>% summarise(mean(pct_male), mean(pop_density), mean(pct_white))

# Transform for facets
best <- best %>% gather(key = variable, Value = pct_male:pct_white)

ggplot(best,aes(x=State, y=value)) + geom_bar(stat='identity') + facet_grid(variable~., scales = "free")  + labs(title = "Best States in 2012", x='',y='')


```

So, there are some possible indicators viewed this way. Racial mix, in the end seems, unlikely. However, larger, less policed areas with less people may be a contributing factor, and a greater percentage of males may increase the odds of one being prone to misconduct. A perhaps quick way to summarize them is below. 

```{r}
# make a table of averages
s <- c("Worst","Best")
sumt <- rbind(wsum,bsum)
sum <- data.frame(s,sumt)
kable(sum)

```




