---
title: "Dougs Crime Analysis"
author: "Doug Walton"
date: "April 26, 2016"
output: html_document
---

This document contains the code I am using to manipulate the demographic and crime files.

Some focusing questions:
1. Are demographics correlaed with the prevalance of either homicide or rape?
Possible factors

	1. household type
	2. age
	3. sex
	4. racial mix (white, black, hispanic)
	5. education
	6. individual income (median)

Approach #1
Starting with U.S. census bureau data to put together a working table for national population and looking a historical approach - how things changed over time. 

Approach #2
Comparing state-to-state and county-to-county.


BUILD UP of APPROACH 1

Starting with tables derived from the 2012 Statistical Abstract of the United States http://www.census.gov/library/publications/2011/compendia/statab/131ed.html

1. us_national.csv is population by year from 1900 to 2015. initially derived from Section 1, Tables 1 & 2. Lagging zeros filled in on earlier years to make everything in consistent units (removed '000s reduction). Added years 2010 to 2015 from PEPANNRES annual estimates of population (found through American Fact Finder.)

I then take selected chosen areas from different census tables to add to this national master.

2. Households in the US. According to the census data dictionary: "A household that has at least one member of the household related to the householder by birth, marriage, or adoption is a "Family household."  Same-sex couple households are included in the family households category if there is at least one additional person related to the householder by birth or adoption.  Same-sex couple households with no relatives of the householder present are tabulated in nonfamily households.  Responses of "same-sex spouse" were edited during processing to "unmarried partner."  "Nonfamily households" consist of people living alone and households which do not have any members related to the householder."
file ACS_14_5YR_B11002.xls and B11001, B11001A, B11001I,used to
B11001
3. Income. This table gives income in 2009 dollars for varous ethnic groups and males/females. Since, the ways that income has has been reported seems to have shifted in 2002 such that some categories like "White" have subgroups instead, some adjustments were made. The subgroups are similar, so I just averaged them back into the top line. Those distinctions aren't going to make much of a difference for a modeling the fit (or could be more fully explored later).

4. Educational attainment. Only high school+ and college+ were available.

5. Compiled from .... and also UCR ucr-us-murder1960-2012.csv. , ucr-us-rape1960-2012.csv


```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(readr)

# Construction of master table for US national data by year
uspop <- read.csv("us_national.csv")
uspop$year <- as.character(uspop$year)

# Add household data - for a test will see if families in couples has any correpondence
UShousehold <- read.csv("Sources/12s0059-households.csv")
UShousehold$Type.of.unit <- as.character(UShousehold$Type.of.unit)
UShousehold$Type.of.unit[3:9] <- c("Household_white","Household_black","Household_hispanic","fam_household","couple_household","male_household","fem_household")
UShousehold_s <- UShousehold %>% slice(1:9) %>% select(-X1960) # eliminate gap
cols <- UShousehold_s$Type.of.unit
UShousehold_s <- as.data.frame(t(UShousehold_s[,-1]))
colnames(UShousehold_s) <- cols
UShousehold_s <- add_rownames(UShousehold_s, "year")
# UShousehold$year <- as.factor(UShousehold$year)
usnat <- uspop %>% left_join(UShousehold_s, by = 'year')

# Income
USincome <- read.csv("Sources/12s0701-median-income-2009-dollars.csv")
USincome$X <- as.character(USincome$X)
cols <- USincome$X
USincome_s <- as.data.frame(t(USincome[,-1]))
colnames(USincome_s) <- cols
USincome_s <- add_rownames(USincome_s, "year")
usnat <- usnat %>% left_join(USincome_s, by = 'year')

# Educational Attainment
USedu <- read.csv("Sources/12s0230-education.csv")
USedu$year <- as.character(USedu$year)
usnat <- usnat %>% left_join(USedu, by = 'year')

# population, age, sex 2000 to 2010
USbasic2010 <- read.csv("Sources/US-EST00INT-ALLDATA.csv")
USbasic2010 <- USbasic2010 %>% filter(AGE != '999' & MONTH != '4')
USbasic2010 <- USbasic2010[,-1]
USbasic2010sum <- USbasic2010 %>% group_by(YEAR) %>% summarise(avg_age = mean(AGE), tot_male = sum(TOT_MALE), tot_female = sum(TOT_FEMALE), white_male_pop = sum(WA_MALE),white_female_pop = sum(WA_FEMALE),black_male_pop = sum(BA_MALE),black_female_pop = sum(BA_FEMALE),hisp_male_pop = sum(H_MALE),hisp_female_pop = sum(H_FEMALE))
USbasic2010sum <- rename(USbasic2010sum, year = YEAR)
USbasic2010sum$year <- paste0("X",USbasic2010sum$year)
usnat <- usnat %>% left_join(USbasic2010sum, by = 'year')

# Adding crime data
UScrime <- read.csv("Sources/12s0306-tot_in_thousands_ofcrimetype1980-2009.csv")
UScrime$year <- paste0("X",UScrime$year)
UScrime[2:6] <- UScrime[2:6] * 1000
usnat2 <- usnat %>% left_join(UScrime, by = 'year')



```

Households. First, we explore the question of whether perhaps more "broken" or non-traditional male-female households could create an increase in crime. So for an indicator, the relationship of households in couples is compared against the the homicide and rape crime rates. 

```{r}
# making a special table, households was in 1000s
usnat_house_murder <- usnat2 %>% filter(year <= 'X2012' & year >= 'X1970') %>% mutate(murderpct = murder/residents, housepct = (couple_household/residents)*1000)



usnat_house_murder  %>% ggplot(aes(year,murderpct)) + geom_point() 

+ geom_point(aes(y = housepct))

# probably doesn't mean anything
cor(usnat3$crimepct,usnat3$housepct)
cor.test(usnat3$crimepct,usnat3$housepct)


```

Notes: So while a historical perspective is possible using a wide varity of data at the national level, it is not clear how to get it at a state, county, or city level without tediously compiling it from many separate files. 


Now we go on to trying the UScensus2010 package. This bit required some tricky installation on my machine. I don't think there is a need to download blk, blkgrp, county, or cd unless or until creating maps. 

a. Seems to be mostly for spatial/visualization work, but has demographics pull function that might be easier than making tables with American Fact Finder. 

b. is only for a deep or comparative slice in certain year essentially (such as the 2010 census)




```{r}
library(UScensus2010)
library(UScensus2010tract)
data("california.tract10")
names(california.tract10)

# msa is the only one that is dem data, the others are spatial data
sfDem <- demographics(dem = c("P0010001", "P0030002", "P0030003"), "CA", level = "msa", msaname = "San Francisco")
colnames(sfDem) <- c("Total Pop","White","Black")
sfDem

# I write this as a csv so you can see the result without using the package
write.csv(sfDem, file = "sfDem-example.csv")
```



Source: Uniform Crime Reporting database (FBI). Has a table builder and can do years by state from 1960 on, so could be interesting. 
http://www.ucrdatatool.gov/index.cfm

Notes:
Can trace down to the LE agency level, but it would be a lot of work to match them automatically to county level stats from the census. 


```{r}




```











Source: U.S. Bureau of Justice Statistics
Covers many aspects of the justice system. The reported crimes focus on the type of vicitimization. Time period covered is about 1993 onward for types of victimization. 

http://www.bjs.gov/

```{r}



```





This is just holding spot for some archive or work in progress

```{r, echo=FALSE}
# some archival stuff
# Add household data - old working
UShousehold <- read.csv("Sources/12s0059-households.csv")
cols <- UShousehold$Type.of.unit
UShousehold <- UShousehold[,-1]
UShousehold <- as.data.frame(t(UShousehold))
colnames(UShousehold) <- cols
UShousehold <- add_rownames(UShousehold, "year")
USnat <- USnat %>% left_join(UShousehold, by = 'year')

# Function for US Census dataframe transpose - nonfunctional at the moment
dftrans <- function(dat) {
  cols <- UShousehold$Type.of.unit
  dat <- as.data.frame(t(dat[,-1]))
  colnames(dat) <- cols
  dat <- add_rownames(dat, "year")
  return(dat)
}

# method for using crime victimization types 
UScrime <- read.csv("Sources/12s0315-crime.csv")
UScrime$X <- as.character(UScrime$X)

#Clean up and filter out some key ones
UScrime_s <- UScrime %>% filter(X %in% c("All_crimes_total","Crimes of violence","..Rape/sexual assault","..Robbery","..Assault","Property_crimes_total","Theft"))

#Prepare the df
cols <- UScrime_s$X
UScrime_s <- as.data.frame(t(UScrime_s[,-1]))
colnames(UScrime_s) <- cols
UScrime_s <- add_rownames(UScrime_s, "year")
# UShousehold$year <- as.factor(UShousehold$year)
usnat2 <- usnat %>% left_join(UScrime_s, by = 'year')


```

Process Notes
1. there are a bunch of footnotes and assumptions made, would probably need to re-examine the effect at some point. 
2. There are fields selected from tables, maybe others were important.
3. fields and ways of accounting for things changing over time

